---
title: "Amphipod Tube Beds as Ecological Phenomena in False Bay: Size Class 
Distribution of Ampelisca spp."
author: "Muhammad Uzair Davids"
date: "2025-03-28"
version: "R version 4.4.3 (2025-02-28 ucrt) -- 'Trophy Case'"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## **1. Mapping the Sampling Stations**
### **1.1. Loading in packages**

I start by loading in the packages I will use to create my map.

```{r}
library(sf) # This package will allow us to read shapefiles.
```

### **1.2. Reading in the data**

The shapefiles I am using to create this map can be downloaded as one .zip file 
[here](https://www.dropbox.com/scl/fi/9fz9bgilr2pp98uoyxthl/Shapefiles_False_Bay.zip?rlkey=msl4u8sleeho5x8ftxyttsp5w&st=rdzdgjw2&dl=0). Note that the .zip file is very large (793.29 MB).

I will now read the data from a data file on my personal computer using the `st_read()` function.

> NOTE: When you are writing/reading this code from your personal computer, you will need to edit the file path to reflect where you have stored your own data. You will need to do this each time I read in data or save data to a file on my personal computer. The `setwd()` function can help you to do this. Alternatively, you can use the `getwd()` function to see what your current working directory is and manually edit the code based on that.

```{r}
# South African province shapefile
province <- st_read("C:/Users/Dell/OneDrive/Desktop/UCT/Honours Project/Ameplisca-Size-Class-Distribution/Data/Shapefiles_False_Bay/SA Provinces/provincial.shp") 

# Coastline shapefile
coast <- st_read("C:/Users/Dell/OneDrive/Desktop/UCT/Honours Project/Ameplisca-Size-Class-Distribution/Data/Shapefiles_False_Bay/SA Coastline/SA Coastline.shp") 

# Depth strata shapefile
depth_strata <- st_read("C:/Users/Dell/OneDrive/Desktop/UCT/Honours Project/Ameplisca-Size-Class-Distribution/Data/Shapefiles_False_Bay/SA Depth Strata - NSBA/depth_strata.shp") 

# World Depth Contour shapefile
depth_contour <- st_read("C:/Users/Dell/OneDrive/Desktop/UCT/Honours Project/Ameplisca-Size-Class-Distribution/Data/Shapefiles_False_Bay/World Depth Contours - hires - 2019/gebco_2019_contours.shp")
```


