---
title: "Size Class Distribution"
author: "Muhammad Uzair Davids"
format: html
editor: visual
---

## 1. Setting Up the Analysis Environment

```{r, echo = FALSE}
# Packages for data wrangling
library(tidyverse)
library(dplyr)

# Packages for creating plots
library(ggplot2)
library(gridExtra) # To arrange plots
library(patchwork) # To arrange plots
```

## 2. Reading the Data

```{r}
# Read the data from the directory
ampelisca <- read.csv("C:/Users/Dell/OneDrive/Desktop/UCT/Honours Project/Ameplisca-Size-Class-Distribution/Data/data_size.csv")

# Let's have a look at the data
summary(ampelisca)
```

## 3. Tidying Up

These column names are a bit of a mess, so let's neaten them up a bit using the `dplyr` package.

```{r}
# Rename columns
ampelisca <- ampelisca %>%
  rename(
    site = Site, # This is just for the sake of consistency
    season = Season, # Same here
    size_adult = Size..mm., # The size of adults in mm
    egg_presence = Eggs..Y.N., # Whether or not adults are carrying eggs (Y/N)
    egg_count = Number.of.Eggs, # Number of eggs the adult is carrying
    juv_presence = Juveniles..Y.N., # Whether or not adults are carrying juveniles (Y/N)
    juv_count = Number.of.Juveniles, # Number of juveniles the adult is carrying
    size_juv = Size.of.Juveniles..mm. # The size of the juveniles being carried in mm
  )
```

## 4. Sorting the Data

Next, I need to separate the data into groups based on their site and season.

```{r}
# Split the data into separate data frames based on season
summer_data <- ampelisca %>% filter(site == "FB005A", season == "Summer")
winter_data <- ampelisca %>% filter(site == "FB005A", season == "Winter")
autumn_data <- ampelisca %>% filter(site == "FB005A", season == "Autumn")
spring_data <- ampelisca %>% filter(site == "FB005A", season == "Spring")
```

We can check to see if it worked by printing the results. If it worked correctly, there should be 100 rows in the new data frames.

```{r}
print(summer_data)
```

Hooray! It worked. Let's do some basic summaries of the data now.

## 5. Counts

The data set contains information regarding how adult *Ampelisca* were carrying eggs and juveniles, as well as how many eggs and juveniles these adults were carrying. We count these using `dplyr`.

Let's start with counting how many adults were carrying eggs in each season.

```{r}
# Winter data
eggs_pres_win <- winter_data %>%
  group_by(egg_presence) %>%
  summarise(count = n())

# Spring data
eggs_pres_spr <- spring_data %>%
  group_by(egg_presence) %>%
  summarise(count = n())

# Summer data
eggs_pres_sum <- summer_data %>%
  group_by(egg_presence) %>%
  summarise(count = n())

# Autumn data
eggs_pres_aut <- autumn_data %>%
  group_by(egg_presence) %>%
  summarise(count = n())

# Print the summaries
print(eggs_pres_win)
```

> Note: The number of adults carrying eggs is recorded in the `Y` row here. Also, the total number of individuals in the `N` and `Y` column should sum to 100.

But how many eggs are present in total?

```{r}
# Summarize: sum of egg_count per season
eggs_per_season <- ampelisca %>%
  group_by(season) %>%
  summarise(total_eggs = sum(egg_count, na.rm = TRUE))  # na.rm = TRUE ignores missing values

# Reorder the 'season' factor levels tp match the order they were sampled in
eggs_per_season$season <- factor(eggs_per_season$season, 
                                 levels = c("Winter", "Spring", "Summer", "Autumn"))

# View the result
print(eggs_per_season)
```

Let's plot the number of eggs per season using `ggplot2`.

```{r}
# Plot of the number of eggs per season
ggplot(eggs_per_season, aes(x = season, y = total_eggs, fill = season)) +
  geom_col() +
  labs(title = "Total Number of Eggs per Season", x = "Season", y = "Total Eggs") +
  theme_minimal() +
  scale_fill_brewer(palette = "Set3")
```

Now let's check how many adults were carrying juveniles in each season - but tomorrow, I'm tired now.

```{r}
# Summarize counts for juv_presence
juvenile_presence <- ampelisca %>%
  group_by(juv_presence) %>%
  summarise(count = n())

print(juvenile_presence)
```

## 6. Plotting Histograms

Let's plot some histograms of the size classes of adults across seasons. But first, I need to know the minimum and maximum sizes of adults so we can set a consistent scale.

```{r}
range(ampelisca$size_adult, na.rm = TRUE)
```

So, the smallest amphipod in our dataset is 2.102mm,, and the largest amphipod is 13.345mm. We can use this range to define the upper and lower limits of our histograms.

```{r}
limits <- c(2.102, 13.345)
```

Now to actually plot the histograms.

```{r}
# Winter
p_winter <- ggplot(winter_data, aes(x = size_adult)) +
  geom_histogram(binwidth = 1, fill = "lightblue", color = "black") +
  ylab("Count") +
  xlab("Size") +
  scale_x_continuous(labels = function(x) paste0(x, " mm"), limits = common_limits) +
  ggtitle("Winter") +
  theme_minimal()

# Spring
p_spring <- ggplot(spring_data, aes(x = size_adult)) +
  geom_histogram(binwidth = 1, fill = "green", color = "black") +
  ylab("Count") +
  xlab("Size") +
  scale_x_continuous(labels = function(x) paste0(x, " mm"), limits = common_limits) +
  ggtitle("Spring") +
  theme_minimal()

# Summer
p_summer <- ggplot(summer_data, aes(x = size_adult)) +
  geom_histogram(binwidth = 1, fill = "orange", color = "black") +
  ylab("Count") +
  xlab("Size") +
  scale_x_continuous(labels = function(x) paste0(x, " mm"), limits = common_limits) +
  ggtitle("Summer") +
  theme_minimal()

# Autumn
p_autumn <- ggplot(autumn_data, aes(x = size_adult)) +
  geom_histogram(binwidth = 1, fill = "brown", color = "black") +
  ylab("Count") +
  xlab("Size") +
  scale_x_continuous(labels = function(x) paste0(x, " mm"), limits = common_limits) +
  ggtitle("Autumn") +
  theme_minimal()

# Combine
grid.arrange(p_winter, p_spring, p_summer, p_autumn, ncol = 2)
```

Great, but I want publication ready plots, so no colours or grids.

```{r}
clean_theme <- theme_minimal(base_size = 12) +
  theme(
    plot.title = element_text(hjust = 0.5, face = "bold"),
    axis.title = element_text(face = "bold"),
    panel.grid.major = element_blank(),   # remove major grid lines
    panel.grid.minor = element_blank(),   # remove minor grid lines
    panel.border = element_blank(),       # remove borders
    axis.line = element_line(color = "black")  # keep axis lines for clarity
  )


make_hist <- function(ampelisca, title) {
  ggplot(ampelisca, aes(x = size_adult)) +
    geom_histogram(binwidth = 1, fill = "grey70", color = "black") +
    xlab("Size (mm)") +
    ylab("Count") +
    scale_x_continuous(limits = common_limits) +
    ggtitle(title) +
    clean_theme
}

# Generate the four histograms in greyscale
p_winter  <- make_hist(winter_data,  "Winter")
p_spring  <- make_hist(spring_data,  "Spring")
p_summer  <- make_hist(summer_data,  "Summer")
p_autumn  <- make_hist(autumn_data,  "Autumn")

# Combine and display
grid.arrange(p_winter, p_spring, p_summer, p_autumn, ncol = 2)
```

I have no idea why this isn't working, I'll try it again later.

## 7. Kolmogorov-Smirnov Test

Let's play around with the `ks.test()` function.

> Note: This is not the way I would actually conduct the test for my final, at this point I am just getting a feel for the function.

A test for normality.

```{r}
# Example: K-S test for normality of 'size' column
ks.test(ampelisca$size_adult, "pnorm", mean = mean(ampelisca$size_adult), sd = sd(ampelisca$size_adult))

```

Comparing size between summer and winter.

```{r}
# Two-sample K-S test
ks.test(summer_data$size_adult, winter_data$size_adult)
```
